MSc Thesis

### 2023-05-26 manuscript/main.tex
added the experimental schematic

To test the prevalence of rapid silicon accumulation in canadian cereal crops, we selected three cultivars for each of oats, bread wheat, triticale, and barley. We selected cultivars on the basis of minimizing shared pedigree, and no cultivars shared more than one common ancestor within the last two crossing generations. At the start of the experiment, we germinated seeds in germination trays filled with moist sand. After four days, we transplanted germinated seedlings into 10 cm pots filled with SunGro potting mix amended with 0.17 g of silicic acid (a rate of 2.4 g silicon /kg of soil). Though potting mix and fresh water contain some amount of plant available silicon, we added the silicic acid to ensure that there would be minimal silicon limitation to the plants. We randomized the location of each pot within the growing space. A flood table bottom watered the pots with nutrient solution. We assigned each plant to one of three herbivory treatments: control, simulated herbivory, or true herbivory. We simulated herbivory by application of 1 mM methyl-jasmonate solution to the entire above-ground portion of the plant (Waterman et al. 2021b), while crickets housed in water-pik tubes provided true herbivory. Prior to introduction to the the plants, we acclimated crickets by feeding them for 72 hours on the same species used in this trial. Immediately preceding cricket application, we placed them in their tubes and starved them for 24 hours, as this increased the likelihood of the insects initiating feeding rapidly upon exposure to the test plants. Prior to harvest, we recorded whether the crickets had initiated feeding on the plants by visually inspecting the leaves for missing tissue.

@@

Recent research in inducible silicon plant defenses has focussed on the short-term dynamics of silicon uptake (\cite{waterman_short-term_2021,waterman_short-term_2021-1}). The promising results of this work has been highlighted for its potential applications in agriculture, where sensitive and rapid defensive phenotypes could improve plant performance and reduce reliance on more intensive pest-control measures. In this study, we attempted to demonstrate rapid silicification in four cereal crops. We failed to find evidence of rapidly induced silicon uptake in response to either methyl-jasmonate application or herbivore exposure.

@@

>>>>

To test the prevalence of rapid silicon accumulation in canadian cereal crops, we selected three cultivars for each of oats, bread wheat, triticale, and barley. We selected cultivars on the basis of minimizing shared pedigree, and no cultivars shared more than one common ancestor within the last two crossing generations. At the start of the experiment, we germinated seeds in germination trays filled with moist sand. After four days, we transplanted germinated seedlings into 10 cm pots filled with SunGro potting mix amended with 0.17 g of silicic acid (a rate of 2.4 g silicon /kg of soil). Though potting mix and fresh water contain some amount of plant available silicon, we added the silicic acid to reduce silicon limitation to the plants. We randomized the location of each pot within the growing space. A flood table bottom watered the pots with nutrient solution. We assigned each plant to one of three herbivory treatments: control, simulated herbivory, or true herbivory. We simulated herbivory by application of 1 mM methyl-jasmonate solution to the entire above-ground portion of the plant (Waterman et al. 2021b), while crickets housed in water-pik tubes provided true herbivory. Prior to introduction to the the plants, we acclimated crickets by feeding them for 72 hours on the same species used in this trial. We applied our treatments 45 days after transplanting. Immediately preceding cricket application, we placed them in their tubes and starved them for 24 hours, as this increased the likelihood of the insects initiating feeding rapidly upon exposure to the test plants. Prior to harvest, we recorded whether the crickets had initiated feeding on the plants by visually inspecting the leaves for missing tissue.

@@

Recent research in inducible silicon plant defenses has focussed on the short-term dynamics of silicon uptake (\cite{waterman_short-term_2021,waterman_short-term_2021-1}). The promising results of this work have been highlighted for their potential applications in agriculture, where sensitive and rapid defensive phenotypes could improve plant performance and reduce reliance on more intensive pest-control measures. In this study, we attempted to demonstrate rapid silicification in four cereal crops. In our study, we failed to find evidence of rapidly induced silicon uptake in response to either methyl-jasmonate application or herbivore exposure.

@@

\begin{figure}[ht]
        \includegraphics[width = \textwidth]{images/Induction_schematic.png}
        \centering
        \caption{Experimental schematic for this study. We grew four species of cereal crops (left box) in a glasshouse. After 45 days, we applied one of three treatments to each plant, in a fully randomized manner (centre box). To analyse the response of leaf silicon and phenolic content, we used X-ray fluorescence (XRF) and the Fast Blue BB phenolic assay (right box).}
\end{figure}
